---
title: "Assignment 2: Data Cleaning"
format: html
---

```{r}
#| message: false  

# Libraries
library(tidyverse)

# file names
datadir_raw <- "data/raw/"

datadir_processed <- "data/processed/"

snowsurvey_file <- "snow_cover.csv"
```

## Read in Data
```{r}
# Import the snow cover data with snow_cover already cleaned
snowsurvey_csv <- read_csv(file.path(datadir_processed, snowsurvey_file))

glimpse(snowsurvey_csv)
```

# 1. Cleaning Water Cover Column

Based on the glimpse, water cover is currently a character, so let's see what non-numeric unique values are in the column.

```{r, warning = FALSE}
snowsurvey_csv %>% 
  count(Water_cover) %>% 
  filter(is.na(as.numeric(Water_cover)))
```

Let's convert all of these to NA values, since that is what we also did with snow cover.

```{r}
# All together using case_when
snowsurvey_water_fixed <- snowsurvey_csv %>% 
  mutate(Water_cover = case_when(
    Water_cover == "-" ~ NA,
    Water_cover == "." ~ NA,
    Water_cover == "n/a" ~ NA,
    Water_cover == "unk" ~ NA,
    # everything else keep the same
    .default = Water_cover
  ))
```

## Transform column to numeric (correct data type)
```{r}
snowsurvey_water_fixed <- snowsurvey_water_fixed %>% 
  mutate(Water_cover = as.numeric(Water_cover))
```

## Check for appropriate percentage values
```{r}
snowsurvey_water_fixed %>% 
  filter(Water_cover > 100)

snowsurvey_water_fixed %>% 
  filter(Water_cover < 0)
```

There is one row where water cover is listed as 353. We can't infer what the number was supposed to be since land cover is listed as -298, so I am going to convert all cover values to NA since I want to keep the record of this observation.
```{r}
snowsurvey_water_fixed <- snowsurvey_water_fixed %>% 
  mutate(Water_cover = ifelse(Water_cover > 100, NA, Water_cover))
```

There are no values of water cover that are less than 0.



# 2. Cleaning land cover column

Land is currently a character so we will repeat the same steps as water, starting with finding the non-numeric values.

```{r}
snowsurvey_water_fixed %>% 
  count(Land_cover) %>% 
  filter(is.na(as.numeric(Land_cover)))
```

The non-numeric values are similar to those in water cover, so we will convert them to NA values
```{r}

```



## Write to a csv file
```{r}
write_csv(snowsurvey_water_fixed, file.path(datadir_processed, "water_cover.csv"))
```